### 2021-07-25 notebooks/DataCleaning.Rmd
started datacleaning file

---
title: "Lab2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r import libraries}
library(tidyverse)
library(ggplot2) 
library(sandwich)
library(stargazer)

```

```{r read in data}

d_covid <- read.csv(file = './JAL_Lab2/src/data/covid-counties.csv')
d_openings <- read.csv(file = './JAL_Lab2/src/data/COVID-19 US state policy database (CUSP) - Closures & Reopening.csv')
d_trips <- read.csv(file = './JAL_Lab2/src/data/Trips_by_Distance_-_2020.csv')
```