### 2021-07-27 notebooks/lab2_report.Rmd
Initial report commit.

---
title: 'Title!'
author: "Autumn Rains, Jonas Degnan, Lucy Wu"
date: "8/5/2021"
output:
  bookdown::pdf_document2: 
    toc: true
---

\clearpage

```{r install packages, echo=FALSE, warning=FALSE, message=FALSE}
# install.packages('tinytex')
# tinytex::install_tinytex()
```

```{r load packages, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse) 
library(magrittr)
library(knitr)
library(patchwork)
library(kableExtra)
library(tinytex)

theme_set(theme_bw())

options(tinytex.verbose = TRUE)
```

```{r load clean data, echo=FALSE, message=FALSE}

```