### 2021-08-03 notebooks/accounting_table.Rmd
added unemployment data

---
title: "accounting table"
output: html_document
---

```{r load packages}
library(dplyr)
library(tinytex)
library(knitr)
library(kableExtra)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r accounting-data}
stage <- c("raw", "date_state", "county", "model_1", "model_2", "model_3")
covid <- c(1576181, 7620, 254,230,"","")
mobility <- c(97402, 7620, 254,230,"","")
urb_rural <- c(3220, "value_2","","","","")
econ <- c(3220, , )
income <- c("value_1", "value_2","","","","")
adjacency <- c("value_1", "value_2","","","","")

df <- data.frame(stage, covid, mobility, urb_rural, income, adjacency)

```

``` {r make population summary table, echo=FALSE}
accounting_summary <- table(
    df$stage, 
    df$covid),
    df$mobility,
    df$urb_rural,
    df$income,
    df$adjacency
)
# names(dimnames(party_summary)) <- c("Self-reported", "Sentiment-revealed")
```

``` {r accounting-table, echo=FALSE}
kbl(accounting_summary, 
    digits = 2,
    caption = 'Accounting Table for Data Source Cleaning',
    booktabs = TRUE) %>%
  kable_paper(c('hover', 'striped'), full_width = T)
```