### 2021-08-03 src/data/DataCleaning.Rmd

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r import libraries, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
# source(local=TRUE, file="../util.R")
```

```{r read in data, echo=FALSE}
d_covid <- read.csv(url('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv'))
d_openings <- read.csv(file = './../../data/raw/COVID-19 US state policy database (CUSP) - Closures & Reopening.csv')
d_trips <- read.csv(file = './../../data/raw/Trips_by_Distance.csv')
```

```{r filter parameters, echo=FALSE}
# state(s) under investigation
state_id = c("TX","Texas",48)

# dates under investigation
start_date = as.Date("2021-06-01")
end_date = as.Date("2021-06-30")

# start_date = as.Date("2021-06-01")
# end_date = as.Date("2021-06-30")
```

```{r clean covid data, echo=FALSE}
# COVID-19 data
# filters by date/state
d_covid <- d_covid %>%
  mutate(
    date = as.Date(date)
  ) %>%
  filter(
    (date >= start_date & date <= end_date),
    grepl(paste(state_id, collapse = "|"), state)
  )

# max county data (cases and deaths are cumulative)
d_covid <- d_covid %>%
  group_by(fips) %>%
  mutate(
    cases = max(cases) - min(cases),
    deaths = max(deaths) - min(deaths),
    state = state.abb[match(state, state.name)]
  ) %>%
  select(
    county, state, fips, cases, deaths
  ) %>%
  distinct(fips, .keep_all = TRUE)
```

```{r clean mobility data, echo=FALSE}
# clean mobility data

# filters by date/state
d_trips <- d_trips %>%
  mutate(
    Date = as.Date(Date)
  ) %>%
  filter(
    (Date >= start_date & Date <= end_date),
    grepl(paste(state_id, collapse = "|"), State.Postal.Code)
  )

# cleaning
d_trips <- d_trips %>%
  # clean county name format
  mutate(
    County.Name = gsub("([A-Za-z]+).*", "\\1", County.Name),
    total_pop = select(.,Population.Staying.at.Home:Population.Not.Staying.at.Home) %>% rowSums(na.rm = TRUE)
    ) %>%
  group_by(County.FIPS) %>%
  mutate(
    # define long trips
    long_trip = sum(c_across(Number.of.Trips.100.250:Number.of.Trips...500)),
    long_trip_pc = long_trip/total_pop,
    # define short trips
    short_trip = sum(c_across(Number.of.Trips.1.3:Number.of.Trips.50.100)),
    short_trip_pc = short_trip/total_pop
  ) %>%
  rename(
    state = State.Postal.Code,
    county = County.Name,
    fips = County.FIPS
  ) %>%
  select(
    county, fips, state, total_pop, long_trip, long_trip_pc, short_trip, short_trip_pc
  ) %>%
  distinct(fips, .keep_all = TRUE)
```

```{r merged model_1 dataframe, echo=FALSE}
#code to merge county info to d
d <- merge(x = d_trips, y = d_covid, by = c("fips", "county", "state"))

#drop na and duplicates
d <- d %>% drop_na()
d <- unique(d)
```

# Clean other variables
```{r read in variable data, echo=FALSE}
density <- read.delim("./../../data/raw/2019_Gaz_counties_national.txt", header=TRUE, sep="\t")
dp03_econ <- read.csv(file = './../../data/raw/ACSDP5Y2019.DP03_data_with_overlays_2021-07-26T143523.csv')
dp05_age <- read.csv(file = './../../data/raw/ACSDP1Y2019.DP05_data_with_overlays_2021-07-21T215013.csv')
s1901_income <- read.csv(file = './../../data/raw/ACSST1Y2019.S1901_data_with_overlays_2021-07-21T222906.csv') 
urb_rur_h2 <- read.csv(file = './../../data/raw/DECENNIALSF12010.H2_data_with_overlays_2021-06-08T173848.csv')

```

```{r clean dp03 data, echo=FALSE}
# filters by date/state
econ <- dp03_econ %>%
  rename(fips = GEO_ID, unemp_rate = DP03_0009PE) %>%
  select(fips, unemp_rate) %>%
  mutate(
    fips = substring(fips, 10),
    unemp_rate = as.double(unemp_rate)
  )

# merge into main dataframe
d <- merge(x = econ, y = d, by = c("fips"))
```

```{r clean dp03 data, echo=FALSE}
# filters by date/state
age <- dp05_age %>%
  rename(fips = GEO_ID,
         female = DP05_0003PE, 
         pct_lt5 = DP05_0005PE, 
         pct_5to9 = DP05_0006PE, 
         pct_10to14 = DP05_0007PE, 
         pct_15to19 = DP05_0008PE, 
         pct_20to24 = DP05_0009PE, 
         pct_25to34 = DP05_0010PE, 
         pct_35to44 = DP05_0011PE, 
         pct_45to54 = DP05_0012PE
) %>% rowwise() %>%
  mutate(
    fips = substring(fips, 10),
    female = as.double(female), 
    pct_lt5 = as.double(pct_lt5), 
    pct_5to9 = as.double(pct_5to9), 
    pct_10to14 = as.double(pct_10to14), 
    pct_15to19 = as.double(pct_15to19), 
    pct_20to24 = as.double(pct_20to24), 
    pct_25to34 = as.double(pct_25to34), 
    pct_35to44 = as.double(pct_35to44), 
    pct_45to54 = as.double(pct_45to54),
    pct_gt55 = 100 - sum(c(pct_lt5, pct_5to9, pct_10to14, pct_15to19, pct_20to24, pct_25to34, pct_35to44, pct_45to54)),
    pct_lt19 = sum(c(pct_lt5, pct_5to9, pct_10to14, pct_15to19)),
    pct_19to54 = 100 - pct_gt55 - pct_lt19
    ) %>%
  select(fips, pct_lt19, pct_19to54, pct_gt55)

# merge into main dataframe
d <- merge(x = age, y = d, by = c("fips"))
```

```{r clean s1901 data, echo=FALSE}
# filters by date/state
income <- s1901_income %>%
  rename(fips = GEO_ID,
  median_income = S1901_C01_012E,  
  pct_50to75k = S1901_C01_007E, 
  pct_75to100k = S1901_C01_008E, 
  pct_100to150k = S1901_C01_009E, 
  pct_150to200k = S1901_C01_010E, 
  pct_gt200k = S1901_C01_011E
) %>% rowwise() %>%
  mutate(
    fips = substring(fips, 10),
    median_income = as.double(median_income), 
    pct_50to75k = as.double(pct_50to75k), 
    pct_75to100k = as.double(pct_75to100k), 
    pct_100to150k = as.double(pct_100to150k), 
    pct_150to200k = as.double(pct_150to200k), 
    pct_over_us_median = sum(c(pct_50to75k, pct_75to100k, pct_100to150k, pct_150to200k))) %>%
  select(fips, median_income, pct_over_us_median)

# merge into main dataframe
d <- merge(x = income, y = d, by = c("fips"))
```

# Clean up
```{r remove interim dataframes}
# remove dataframes
rm(d_covid, d_trips, d_openings)
rm(dp03_econ, dp05_age, density, s1901_income, urb_rur_h2, econ, income, age)

```

# Write to .csv
```{r export interim model_1 dataframe as csv, echo=FALSE}
write.csv(d, '../../data/interim/covid_trips_interim.csv')
```