### 2021-08-04 src/data/DataCleaning.Rmd

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r import libraries, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
```

# limited model data cleaning

```{r read in data, echo=FALSE}
d_covid <- read.csv(url('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv'))
d_trips <- read.csv(file = './../../data/raw/Trips_by_Distance.csv')
urb_rur_h2 <- read.csv(file = './../../data/raw/DECENNIALSF12010.H2_data_with_overlays_2021-06-08T173848.csv') 
density <- read.csv(file='./../../data/raw/TAC_population.csv')
income <- read.csv('./../../data/raw/TAC_income.csv')
```

```{r filter parameters, echo=FALSE}
# state(s) under investigation
state_id = c("TX","Texas",48)

# dates under investigation
start_date = as.Date("2021-06-01")
end_date = as.Date("2021-06-30")
```

```{r clean covid data, echo=FALSE}
# COVID-19 data
# filters by date/state
d_covid <- d_covid %>%
  mutate(
    date = as.Date(date)
  ) %>%
  filter(
    (date >= start_date & date <= end_date),
    grepl(paste(state_id, collapse = "|"), state)
  )

# max county data (cases and deaths are cumulative)
d_covid <- d_covid %>%
  group_by(fips) %>%
  mutate(
    cases = max(cases) - min(cases),
    deaths = max(deaths) - min(deaths),
    state = state.abb[match(state, state.name)]
  ) %>%
  select(
    county, state, fips, cases, deaths
  ) %>%
  distinct(fips, .keep_all = TRUE)
```

```{r clean mobility data, echo=FALSE}
# clean mobility data

# filters by date/state
d_trips <- d_trips %>%
  mutate(
    Date = as.Date(Date)
  ) %>%
  filter(
    (Date >= start_date & Date <= end_date),
    grepl(paste(state_id, collapse = "|"), State.Postal.Code)
  )

# cleaning/county filter
d_trips <- d_trips %>%
  # clean county name format
  mutate(
    County.Name = gsub("([A-Za-z]+).*", "\\1", County.Name),
    total_pop = select(.,Population.Staying.at.Home:Population.Not.Staying.at.Home) %>% rowSums(na.rm = TRUE)
    ) %>%
  group_by(County.FIPS) %>%
  mutate(
    # define long trips
    long_trip = sum(c_across(Number.of.Trips.100.250:Number.of.Trips...500)),
    # define short trips
    # short_trip = sum(c_across(Number.of.Trips.1.3:Number.of.Trips.50.100)),
  ) %>%
  rename(
    state = State.Postal.Code,
    county = County.Name,
    fips = County.FIPS
  ) %>%
  select(
    county, fips, state, total_pop, long_trip
  ) %>%
  distinct(fips, .keep_all = TRUE)
```

```{r clean urban/rural data}
rural_ind <- urb_rur_h2[-c(1),] %>%
  rename(fips = GEO_ID,
         total_pop = H002001,
         urbanized = H002003,
         rural = H002005) %>%
  select(fips, total_pop, urbanized, rural) %>%
  mutate(
    fips = substring(fips, 10),
    total_pop = as.double(total_pop),
    urbanized = as.double(urbanized),
    rural = as.double(rural),
    is_rural = ifelse((total_pop < 50000 & urbanized == 0), 1, 0)
  ) %>% 
  select(fips, is_rural) %>% drop_na()
```

```{r clean income data}
income_f <- income %>%
  rename(
    county = County,
    median_income = Median.Household.Income,
    povery_rate = X..of.Population.in.Poverty
  ) %>%
  select(
    county, median_income, povery_rate
  ) %>%
  mutate(
    median_income = as.double(gsub('[$,]', '', median_income))
  )

```

```{r clean density data}
density_f <- density %>%
  rename(
    county = County,
    density = X.Population.Density.Per.Sq.Mile
    # total_pop = County.Population,
  ) %>%
  select(
    county, 
    density 
    #total_pop
  ) %>%
  mutate(
    #total_pop = as.double(gsub('[,]', '', total_pop)),
    density = as.double(gsub('[,]', '', density))
  )

```

```{r merge limited model dataframe}
d <- merge(x = d_trips, y = d_covid, by = c("fips", "county", "state")) 
d <- merge(x = rural_ind, y = d, by = c("fips"))
d <- merge(x = income_f, y = d, by = c("county"))
d <- merge(x = density_f, y = d, by = c("county"))
d <- d %>% drop_na()

knitr::kable(head(d))
#rm(d)
```

# Write to .csv
```{r export interim model_1 dataframe as csv, echo=FALSE}
write.csv(d, '../../data/interim/covid_trips_interim_limited.csv')
```

# Model 2 data cleaning

```{r load unemployment data}
unemploy <- read.csv('./../../data/raw/TAC_unemployment.csv')
```

```{r clean unemployment data}
unemploy_f <- unemploy %>%
  rename(
    county = County,
    unempoy_rate = Unemployment.Rate....
  ) %>%
  select(
    county, 
    unempoy_rate 
  )

```

```{r merge model 2 dataframe}
d <- merge(x = unemploy_f, y = d, by = c("county"))
d <- d %>% drop_na()

knitr::kable(head(d))
#rm(d)
```

# Write to .csv
```{r export interim model_1 dataframe as csv, echo=FALSE}
write.csv(d, '../../data/interim/covid_trips_interim_model2.csv')
```

# Model 3 data cleaning

```{r load age data}
age <- read.csv('./../../data/raw/TAC_age.csv')
```

```{r clean unemployment data}
age_f <- age %>%
  rename(
    county = County,
    pct_lt17 = X..Age.17.and.Under
  ) %>% rowwise() %>%
  mutate(
    pct_gt65 = sum(c(X..Age.65.and.Older, X..Age.85.and.Older))
  ) %>%
  select(
    county,
    pct_lt17,
    pct_gt65
  )

```

```{r merge limited model dataframe}
d <- merge(x = age_f, y = d, by = c("county"))
d <- d %>% drop_na()

knitr::kable(head(d))
#rm(d)
```

# Write to .csv
```{r export interim model_1 dataframe as csv, echo=FALSE}
write.csv(d, '../../data/interim/covid_trips_interim_model3.csv')
```

# Clean up
```{r remove interim dataframes}
# remove dataframes
rm(d_covid, d_trips, d_openings)
rm(dp03_econ, dp05_age, density, s1901_income, urb_rur_h2, econ, income, age)

```